---
title: "Places of Interest in Singapore"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# DESCRIPTIVE ANALYSIS
## Loading Dataset
```{r}
library(sf)

# Set Working Directory
setwd("C:/Users/fanfa/Desktop/DSA303/DSA303_Proj/dataset")

# Parks
parks <- st_read('parks-geojson.geojson')
table(st_is_valid(parks))
parks_kml <- st_read("parks-kml.kml")
table(st_is_valid(parks_kml))

# Historic Sites
historicsites <- st_read('historic-sites-geojson.geojson')
table(st_is_valid(historicsites))

# Monuments
monuments <- st_read('monuments-geojson.geojson')
table(st_is_valid(monuments))

# Museums
museums <- st_read('museums-geojson.geojson')
table(st_is_valid(museums))

# Tourist Attractions
attractions <- st_read('TOURISM.shp')
table(st_is_valid(attractions))

# SG Planning Areas
sg_map <- st_read("MP14_PLNG_AREA_WEB_PL.shp") 
```

## Merging Different Datasets

**Important Note:** Upon further research, we have found that the Leaflet package expects all point, line and shape data to be specified in latitude and longitude which will then be automatically projected into EPSG3857 (Web Mercator), a common CRS used in many online maps such as OpenStreetMap (https://rstudio.github.io/leaflet/projections.html). As we intend to present our findings using functions and interactive maps from the Leaflet package, we have decided to utilise EPSG4236 (longlat) as the CRS for our analysis.

```{r}
library(dplyr); library(tmap)

# Attractions (shapefile format)
attractions_omit <- attractions[-73,] # remove the row with 0 for Long and Lat
attractions_longlat <- attractions_omit %>% st_transform(crs = "+proj=longlat")
st_crs(attractions_longlat)$proj4string
CRS = st_crs(attractions_longlat)$proj4string

attractions_data <- attractions_longlat %>%
  mutate(NAME = PAGETITLE, DESCRIPTION = META_DESCR, HYPERLINK = URL_PATH, PHOTOURL = IMAGE_PATH, OPENING_HOURS = OPENING_HO) %>%
  select(NAME, DESCRIPTION, OPENING_HOURS, ADDRESS, POSTALCODE, HYPERLINK, PHOTOURL, geometry)
glimpse(attractions_data)

# Full_join() on datasets parks, museums, monuments and historicsites (geojson format)
historicsites_df <- as.data.frame(historicsites)
combined_json <- full_join(parks, historicsites_df)

monuments_df <- as.data.frame(monuments)
combined_json1 <- full_join(combined_json, monuments_df)

museums_df <-  as.data.frame(museums)
combined_json2 <- full_join(combined_json1, museums_df)
combined_json3 <- st_zm(combined_json2, drop = TRUE)

combined_json3 <- combined_json3 %>% st_transform(crs = CRS)
st_crs(combined_json3)$proj4string

# extract non-spatial data from the 'Description' column
library(tidyverse); library(rvest)
json_attributes <- lapply(X = 1:nrow(combined_json3), 
                     FUN = function(x) {

                       combined_json3 %>% 
                         slice(x) %>%
                         pull(Description) %>%
                         read_html() %>%
                         html_node("table") %>%
                         html_table(header = TRUE, trim = TRUE, dec = ".", fill = TRUE) %>%
                         as_tibble(.name_repair = ~ make.names(c("Attribute", "Value"))) %>% 
                         pivot_wider(names_from = Attribute, values_from = Value)

                     })

json_attr <- combined_json3 %>% bind_cols(bind_rows(json_attributes)) %>%
  select(-Description)

json_data <- json_attr %>%
  mutate(ADDRESS = ADDRESSSTREETNAME, POSTALCODE = ADDRESSPOSTALCODE) %>%
  select(NAME, DESCRIPTION, ADDRESS, POSTALCODE, HYPERLINK, PHOTOURL, geometry)
glimpse(json_data)

# Full join of all datasets
all_data <- full_join(as.data.frame(attractions_data)[, -8], as.data.frame(json_data)[, -7])
all_data$geometry <- c(attractions_data$geometry,json_data$geometry)
glimpse(all_data)
```

## Adding a column called Category with weights
```{r}
library(tmap); library(aspace, quietly = T); library(leaflet); library(sf)

# Creating a new sf data frame with categories
CATEGORY <- c(rep("Attractions", 106), rep("Parks", 350), rep("Historic Sites", 99), rep("Monuments", 72), rep("Museums", 54))

full_data <- cbind(all_data, CATEGORY) %>% as.data.frame() %>% st_as_sf(crs = CRS)

glimpse(full_data)
class(full_data)

# Applying of weights based on Categories (Selected Category = 10, Other Categories = 1)
att_weights <- c(rep(1, nrow(full_data)))
att_weights[full_data$CATEGORY == "Attractions"] <- 10

parks_weights <- c(rep(1, nrow(full_data)))
parks_weights[full_data$CATEGORY == "Parks"] <- 10

monuments_weights <- c(rep(1, nrow(full_data)))
monuments_weights[full_data$CATEGORY == "Monuments"] <- 10

historicsites_weights <- c(rep(1, nrow(full_data)))
historicsites_weights[full_data$CATEGORY == "Historic Sites"] <- 10

museums_weights <- c(rep(1, nrow(full_data)))
museums_weights[full_data$CATEGORY == "Museums"] <- 10

```

## Basemap and Tourist Interactive Map
```{r}
library(sf); library(tmap)

# Basemap for Descriptive Analysis
tmap_mode("view")
base_layer <- tm_shape(full_data) + tm_dots(col = "CATEGORY", palette = "Set1", size = 0.05, popup.vars = c("DESCRIPTION", "OPENING_HOURS", "ADDRESS", "POSTALCODE", "HYPERLINK", "PHOTOURL")) + tm_basemap(server = "OpenStreetMap")

# Tourist Interactive Map with Layers Control and Descriptions
Attractions <- full_data[full_data$CATEGORY == "Attractions",]
Monuments <- full_data[full_data$CATEGORY == "Monuments",]
Museums <- full_data[full_data$CATEGORY == "Museums",]
Historic_Sites <- full_data[full_data$CATEGORY == "Historic Sites",]
Parks <- full_data[full_data$CATEGORY == "Parks",]

base_layer2 <- tm_shape(Attractions) + tm_dots(col = "blue4", size = 0.05) +  tm_shape(Historic_Sites) + tm_dots(col = "darkgreen", size = 0.05) + tm_shape(Monuments) + tm_dots(col = "cornflowerblue", size = 0.05) + tm_shape(Museums) + tm_dots(col = "orange", size = 0.05) +  tm_shape(Parks) + tm_dots(col = "cyan", size = 0.05) + tm_basemap(server = "OpenStreetMap")

tmap_leaflet(base_layer2) %>%
  addLayersControl(overlayGroups = c("Attractions", "Parks", "Museums", "Monuments", "Historic_Sites"), options = layersControlOptions(collapse = F))
```

## Creating Functions for Descriptive Analysis and Weighted Descriptive Analysis
```{r}
# Function for Descriptive Analysis
descriptive_analysis <- function(Data, coords, CRS){
  
  ## Standard Ellipse - central tendency
  mean = mean_centre(id = 1, points = coords)
  median = median_centre(id = 2, points = coords)
  
  ## sdd
  sdd = calc_sdd(id=3, calccentre = T, weighted = F, points = coords)
  sdd_line = sddloc %>% st_as_sf(coords = c("x", "y"), crs = CRS) %>% st_combine() %>% st_cast("LINESTRING")
  std_dist = st_sf(sddatt, geom = st_geometry(sdd_line))
  
  ## sde
  sde = calc_sde(id=4, calccentre = T, weighted = F, points= coords)
  sde_line = sdeloc %>% st_as_sf(coords = c("x", "y"), crs = CRS) %>% st_combine() %>% st_cast("LINESTRING")
  std_ellps = st_sf(sdeatt, geom=st_geometry(sde_line))
  
  ## mean and median
  mean_median = rbind(meanloc,medianloc) %>% data.frame() %>% st_as_sf(coords=c("x","y"), crs=CRS)
  mean_median$id = c("Spatial Mean","Spatial Median")
  
  ## plot
  base_layer <- tm_shape(Data) + tm_dots(col = "CATEGORY", palette = "Set2", size = 0.05, popup.vars = c("DESCRIPTION", "OPENING_HOURS", "ADDRESS", "POSTALCODE", "HYPERLINK", "PHOTOURL")) + tm_basemap(server = "OpenStreetMap")
  
  final_map <- base_layer + tm_shape(std_ellps) + tm_lines(lwd=1) + tm_shape(std_dist) + tm_lines(lwd=2, lty=2, col = "red") + tm_shape(mean_median) + tm_dots(size=0.05, shape="id", col="red")

  ## leaflet plot with radius around spatial median
  tmap_leaflet(final_map) %>%
  addTiles %>%
  addCircles(lng = st_coordinates(mean_median$geometry[2])[1], lat = st_coordinates(mean_median$geometry[2])[2], radius = 2000)
}

# Function for Weighted Descriptive Analysis
weighted_descriptive_analysis <- function(Data, Coords, CRS, Weights){
  ## Standard Ellipse - central tendency
  mean = mean_centre(id = 1, weighted = T, weights = Weights, points = Coords)
  median = median_centre(id = 2, points = Coords)
  
  ## sdd
  sdd = calc_sdd(id=3, calccentre = T, weighted = T, weights = Weights,  points = Coords)
  sdd_line = sddloc %>% st_as_sf(coords = c("x", "y"), crs = CRS) %>% st_combine() %>% st_cast("LINESTRING")
  std_dist = st_sf(sddatt, geom = st_geometry(sdd_line))
  
  ## sde
  sde = calc_sde(id=4, calccentre = T, weighted = T, weights = Weights, points= Coords)
  sde_line = sdeloc %>% st_as_sf(coords = c("x", "y"), crs=CRS) %>% st_combine() %>% st_cast("LINESTRING")
  std_ellps = st_sf(sdeatt, geom=st_geometry(sde_line))
  
  ## mean and median
  mean_median = rbind(meanloc,medianloc) %>% data.frame() %>% st_as_sf(coords=c("x","y"), crs=CRS)
  mean_median$id = c("Spatial Mean","Spatial Median")

  ## plot
  base_layer <- tm_shape(Data) + tm_dots(col = "CATEGORY", palette = "Set2", size = 0.05, popup.vars = c("DESCRIPTION", "OPENING_HOURS", "ADDRESS", "POSTALCODE", "HYPERLINK", "PHOTOURL")) + tm_basemap(server = "OpenStreetMap")
  
  final_map <- base_layer + tm_shape(std_ellps) + tm_lines(lwd=1) + tm_shape(std_dist) + tm_lines(lwd=2, lty=2, col = "red") + tm_shape(mean_median) + tm_dots(size=0.05, shape="id", col="red")
  
  ## leaflet plot with radius around the weighted spatial mean
  tmap_leaflet(final_map) %>%
  addTiles %>%
  addCircles(lng = st_coordinates(mean_median$geometry[1])[1], lat = st_coordinates(mean_median$geometry[1])[2], radius = 2000)
}
```

## Descriptive Analysis of Places of Interests
```{r}
all_coords <- st_coordinates(full_data)

tmap_mode("view")
poi_map <- descriptive_analysis(full_data, all_coords, CRS)
poi_map
```

## Descriptive Analysis of Tourist Attractions
```{r}
# attraction coordinates
att_coords <- full_data[full_data$CATEGORY == "Attractions",] %>% st_coordinates() 

tmap_mode("view")
att_final_map <- descriptive_analysis(Attractions, att_coords, CRS)
att_final_map
```


## Weighted Descriptive Analysis of Tourist Attractions
```{r}
tmap_mode("view")
att_weighted_map <- weighted_descriptive_analysis(Attractions, all_coords, CRS, att_weights)
att_weighted_map
```


## Descriptive Analysis of Parks
```{r}
# parks coordinates
parks_coords <- full_data[full_data$CATEGORY == "Parks",] %>% st_coordinates() 

tmap_mode("view")
parks_map <- descriptive_analysis(Parks, parks_coords, CRS)
parks_map
```

## Weights Descriptive Analysis of Parks
```{r}
tmap_mode("view")
parks_weighted_map <- weighted_descriptive_analysis(Parks, all_coords, CRS, parks_weights)
parks_weighted_map
```

## Descriptive Analysis of historicsites
```{r}
# historicsites coordinates
historicsites_coords <- full_data[full_data$CATEGORY == "Historic Sites",] %>% st_coordinates() 

tmap_mode("view")
historicsites_map <- descriptive_analysis(Historic_Sites, historicsites_coords, CRS)
historicsites_map
```

## Weights Descriptive Analysis of historicsites
```{r}
tmap_mode("view")
historicsites_weighted_map <- weighted_descriptive_analysis(Historic_Sites, all_coords, CRS, historicsites_weights)
historicsites_weighted_map
```

## Descriptive Analysis of monuments
```{r}
# monuments coordinates
monuments_coords <- full_data[full_data$CATEGORY == "Monuments", ] %>% st_coordinates() 

tmap_mode("view")
monuments_map <- descriptive_analysis(Monuments, monuments_coords, CRS)
monuments_map
```

## Weights Descriptive Analysis of monuments
```{r}
tmap_mode("view")
monuments_weighted_map <- weighted_descriptive_analysis(Monuments, all_coords, CRS, monuments_weights)
monuments_weighted_map
```


## Descriptive Analysis of museums
```{r}
# museums coordinates
museums_coords <- full_data[full_data$CATEGORY == "Museums", ] %>% st_coordinates() 

tmap_mode("view")
museums_map <- descriptive_analysis(Museums, museums_coords, CRS)
museums_map
```

## Weights Descriptive Analysis of museums
```{r}
tmap_mode("view")
museums_weighted_map <- weighted_descriptive_analysis(Museums, all_coords, CRS, museums_weights)
museums_weighted_map
```

# Ideal Location for Accomodation 
```{r}
# Loading Hotels dataset
#setwd("C:/Users/Chin/OneDrive/Desktop/DSA303/Project/Project Dataset")
setwd("C:/Users/fanfa/Desktop/DSA303/DSA303_Proj/dataset")
hotels <- st_read("hotel-locations.kml")
st_crs(hotels)$proj4string

# Initial plot of Hotels dataset
tmap_mode("view")
tm_shape(hotels) + tm_dots() + tm_basemap(server = "OpenStreetMap")
```

## Extracting non-spatial data from Hotels dataset
```{r}
# Extracting data from 'Description' column in the kml dataset
library(tidyverse)
library(rvest)
attributes <- lapply(X = 1:nrow(hotels), 
                     FUN = function(x) {

                       hotels %>% 
                         slice(x) %>%
                         pull(Description) %>%
                         read_html() %>%
                         html_node("table") %>%
                         html_table(header = TRUE, trim = TRUE, dec = ".", fill = TRUE) %>%
                         as_tibble(.name_repair = ~ make.names(c("Attribute", "Value"))) %>% 
                         pivot_wider(names_from = Attribute, values_from = Value)

                     })

hotels_attr <- hotels %>% bind_cols(bind_rows(attributes)) %>%
  select(-Description)

hotels_cleaned <- hotels_attr %>% select(NAME, HYPERLINK, POSTALCODE, TOTALROOMS, ADDRESS, geometry)
glimpse(hotels_cleaned)

# hotels basemap
tmap_mode("view")
hotels_base <- tm_shape(hotels_cleaned) + tm_dots() + tm_basemap(server = "OpenStreetMap")
hotels_base
```


## Tourist Interactive Map - Recommended Location of Stay for Tourists
```{r}
# Calculate spatial medians of each Category of places of interest
AttractionsMedian <- median_centre(id = 2, points = att_coords) %>% st_as_sf(coords = c("median.x", "median.y"), crs = "+proj=longlat") %>% st_transform(crs ="+proj=longlat") %>% st_coordinates()
MonumentsMedian <- median_centre(id = 2, points = monuments_coords) %>% st_as_sf(coords = c("median.x", "median.y"), crs = CRS) %>% st_coordinates()
MuseumsMedian <- median_centre(id = 2, points = museums_coords) %>% st_as_sf(coords = c("median.x", "median.y"), crs = CRS) %>% st_coordinates()
HistoricSitesMedian <- median_centre(id = 2, points = historicsites_coords) %>% st_as_sf(coords = c("median.x", "median.y"), crs = CRS) %>% st_coordinates()
ParksMedian <- median_centre(id = 2, points = parks_coords) %>% st_as_sf(coords = c("median.x", "median.y"), crs = CRS) %>% st_coordinates()

# Interactive Map using Leaflet (OpenStreetMap)
tmap_leaflet(hotels_base) %>%
  addTiles() %>%
  addCircles(lng = AttractionsMedian[1], lat = AttractionsMedian[2], radius = 1000, group = "Attractions", popup = "Attractions") %>%
  addCircles(lng = ParksMedian[1], lat = ParksMedian[2], radius = 1000, group = "Parks", popup = "Parks") %>% 
  addCircles(lng = MonumentsMedian[1], lat = MonumentsMedian[2], radius = 1000, group = "Monuments", popup = "Monuments") %>% 
  addCircles(lng = MuseumsMedian[1], lat = MuseumsMedian[2], radius = 1000, group = "Museums", popup = "Museums") %>% 
  addCircles(lng = HistoricSitesMedian[1], lat = HistoricSitesMedian[2], radius = 1000, group = "Historic Sites", popup = "Historic Sites") %>%
  addLayersControl(overlayGroups = c("Attractions", "Parks", "Monuments", "Museums", "Historic Sites"), options = layersControlOptions(collapsed = F))
```

# POINT PATTERN ANALYSIS
## Quadrat Count Analysis
```{r}
library(spatstat, quietly=T); library(GISTools)

# convert to projection
crs_proj <- st_crs(sg_map)$proj4string
poi_proj <- full_data %>% st_transform(crs = crs_proj)

# convert to PPP
all_ppp <- as.ppp(poi_proj)

# change window
sg_blocks <- sg_map %>% as("Spatial")
class(sg_blocks)
sg.owin <- as(sg_blocks, "owin")
Window(all_ppp) = sg.owin

# qc plot
qc <- quadratcount(all_ppp, nx = 10, ny = 5)
plot(qc, lty = 3, main = "Quadrat Count Plot of Places of Interest")

# quadrat test
(quad.test = quadrat.test(all_ppp, nx=10, ny=5))
# Conclusion: Since P-value < 0.001, we can reject H0 at 1% level of Significance. Hence, conclude that sufficient evidence of clustering.
```


## Kernel Density Estimate of Places of Interest
```{r}
all_ppp2 = as.ppp(poi_proj)
all_kde = density(all_ppp2, sigma = bw.diggle(all_ppp2))
plot(all_kde, main = "KDE: Places of Interest")
```

## Clark Evans Test of Places of Interest
```{r}
sg_buff = gBuffer(sg_blocks, width=1000) # width is in meters
sg_owin = as(sg_buff,"owin")
clarkevans.test(all_ppp2, correction="guard", clipregion = sg_owin, alternative="less")
```

## K-Function Analysis of Places of Interest
```{r}
kf.env <- envelope(all_ppp2, Kest, correction = "best", alternative = "less")
plot(kf.env, main = "K-Function Analysis of Places of Interest")
```